Supplement to: Optimal management of a stochastically varying population when policy adjustment is costly

### 2015-08-05 manuscripts/manuscript.Rmd
number supplement with A\d not S\d

output:
  pdf_document:
    template: components/elsarticle.latex
    fig_caption: true
    keep_tex: true
---

@@

To simplify presentation of the results, we will illustrate cases where the
growth parameters are chosen so that the equilibrium biomass for the equivalent
deterministic model without harvesting is `r (pars[1]-1)/pars[2]`; (specifically, $A=$ `r pars[1]` and
$B=$ `r pars[2]` in Eqn \eqref{eq:state_equation}). To characterize environmental variability, we
assume multiplicative shock $Z_t$ is distributed log-normally with
log standard deviation $\sigma_g$ = `r sigma_g`. In addition, we
show cases where $p=$ `r price`, $\delta=$ `r delta` and $c_0=$ `r c0`.
In the supplement we illustrate that the qualitative patterns observed
here are not sensitive to the specific choices of these parameter values.

@@

In the Supplementary Material, we compare our results with more
conventional fisheries economic formulation in which additional costs are
applied to the control variables themselves, as opposed to adjustments to
the controls: $\Pi_4 (N_t,h_t) = p h_t -  c_0 E_t-c_4E_t^2$.  Additional
costs of this form tend to have a smoothing effect on optimal quotas,
but also change the long-term average stock size or quota size that
is optimal.

@@

To compare the impact of penalty functions on optimal management across
the different functional forms, we select penalty cost coefficients
that induce the same reduction in maximum expected NPV. For example,
the dashed vertical line in Figure 2 maps the needed cost coefficient
$c_i$, for each penalty function, such that the fishery is worth `r 100*(1-reduction)`%
of its unconstrained value when optimally managed. To demonstrate that the results
shown here are independent of this choice of calibrating to a `r 100*(1-reduction)`%
reduction, we conduct the same analysis with a 10% and 30% reduction rate as well.
The corresponding figures can be found in the Supplemental Materials.

@@

Figure 3 illustrates how the different forms of adjustment cost can
impact the dynamics of the optimal harvest policy. Corresponding stock
sizes can be seen in the supplement.  Each panel is generated against the
same sequence of environmental variability so that they can be compared
directly.  The harvest policy chosen shows a systematic deviations from
the adjustment cost free optimum, depending on the structure of the cost function.  
In each case, the optimal solution without any adjustment cost is 
shown by the dashed grey line, with
the policy induced by optimization under the given cost structure 
(equivalent to a `r 100*reduction`% reduction in maximum expected NPV)
overlaid in solid blue.

@@

Figure 4 shows these qualitative patterns still hold at when the models are calibrated 
to a range of different reductions in NPV, from 0% to 30%, though naturally the magnitude
of the differences is greatest when the penalty is larger. Note that the supplementary material provide
further evidence that these patterns are independent of the exact reduction shown by providing
results analogous to Figure 3 (A8, A9) and Figure 5 (A10, A11), and Figure 6 (A6).

@@

```{r}
fig5_caption <- paste("Distribution of revenues from fishing and costs paid to adjustment. Plot shows distributions for the linear costs structure ($\\Pi_1$) only, alternate penalties can be found in the supplementary materials. Dockside revenues are always higher than adjustment costs.  When those costs are not accounted for in the policy ('no adjustment costs'), it is possible to obtain only marginally higher revenues, but pay higher adjustment costs.")
```

@@

```{r Figure_6, fig.cap="Relative cost of ignoring adjustment costs when they are present ('ignoring', red) vs assuming the adjustment costs when they are absent ('assuming', blue).  All values are relative to the cost free adjustment optimum, $NPV_0$.  In each case, it is less costly to assume adjustment costs are present when they are not than it is to ignore them when they are present.Supplementary Figure A6 shows how the magnitude of this effect changes with the size of the reduction to $NPV$.", dev=c("pdf", "svg", "png"), fig.width=7, cache=FALSE}

@@

Figure 6 shows that across each penalty form $\Pi_1$, $\Pi_2$ & $\Pi_3$,
the economic impact of ignoring adjustment costs when they are present
is higher than accounting for them when they are absent. This holds even
after accounting for effect described above: the red bars are consistently
well below the `r (1-reduction)*100`% of $NPV_0$, while the blue bars
are barely below $NPV_0$ (a value of 1 on the vertical axis).  We also
see that the impacts on net present value of the fishery of incorrectly
assuming or ignoring policy adjustment costs are most severe when these
costs are assumed to scale quadratically with the size of the policy
change being implemented ($\Pi_2$).  In the Supplementary Materials we
show that pattern holds across different severity of adjustment costs
(robust to choice of $c_i$), though the larger the adjustment cost,
the greater the difference between these two errors.

Note that this analysis does not consider the possibility that an
adjustment cost is present, but of a different functional form than the
policy assumes.  The consequences of managing under the wrong functional
form can be as or more severe than simply ignoring adjustment costs in
the optimization altogether.  We illustrate several of these mismatch
errors in the Supplementary Material.

@@

The biggest differences between the representations of policy adjustment
cost that we consider are between those that have a smoothing effect on
annual quotas and those that do not. The few past studies that incorporate
policy adjustment costs in models of fisheries or other environmental
management contexts [@Feichtinger1994; @Wirl1999] have assumed costs of
policy adjustment increased as a quadratic function of the magnitude of
the policy change being proposed (analogous to $\Pi_2$ formulation). In
effect, this means that small changes to annual quotas incur little
extra cost, but large changes to annual quotas become disproportionately
expensive to make. Including costs of this form smooths inter-annual
variation in the recommended catch quotas. Also, the catches
and remaining stock sizes corresponding to optimal management become more
autocorrelated in time, because it takes the fishery multiple timesteps to harvest
down peaks in abundance that follow large recruitment pulses. The effects
of smoothing here are similar to those predicted when assuming the cost
per unit effort is increasing in the amount of effort expended [@Brown1974;
@Lewis1981; @McGough2009]; (see Supporting Information for a summary of the
relevant results) as opposed to associating extra costs with changes to
policy *per se*. Smoothing effects of this type are what one might have
expected when including policy adjustment costs.

@@

Assuming an adjustment cost exists when in fact it is absent is not the
same as using the wrong functional form when it is present.  In the supplement
we show that choice of the functional form makes a significant
difference.  We show it can be better to ignore policy costs than to derive a 
policy based on assuming the wrong functional form.  This result further
underscores the importance of modelers using caution when seeking to account
for these costs. Assuming an arbitrary form in order to capture the influence
of adjustment costs is thus unlikely to be instructive.

>>>>

output:
  pdf_document:
    template: components/elsarticle.latex
    fig_caption: true
    keep_tex: true
    includes:
      after_body: appendix_description.tex

---

@@

To simplify presentation of the results, we will illustrate cases where the
growth parameters are chosen so that the equilibrium biomass for the equivalent
deterministic model without harvesting is `r (pars[1]-1)/pars[2]`; (specifically, $A=$ `r pars[1]` and
$B=$ `r pars[2]` in Eqn \eqref{eq:state_equation}). To characterize environmental variability, we
assume multiplicative shock $Z_t$ is distributed log-normally with
log standard deviation $\sigma_g$ = `r sigma_g`. In addition, we
show cases where $p=$ `r price`, $\delta=$ `r delta` and $c_0=$ `r c0`.
In Appendix A we illustrate that the qualitative patterns observed
here are not sensitive to the specific choices of these parameter values.

@@

In Appendix A, we compare our results with more
conventional fisheries economic formulation in which additional costs are
applied to the control variables themselves, as opposed to adjustments to
the controls: $\Pi_4 (N_t,h_t) = p h_t -  c_0 E_t-c_4E_t^2$.  Additional
costs of this form tend to have a smoothing effect on optimal quotas,
but also change the long-term average stock size or quota size that
is optimal.

@@

To compare the impact of penalty functions on optimal management across
the different functional forms, we select penalty cost coefficients
that induce the same reduction in maximum expected NPV. For example,
the dashed vertical line in Figure 2 maps the needed cost coefficient
$c_i$, for each penalty function, such that the fishery is worth `r 100*(1-reduction)`%
of its unconstrained value when optimally managed. To demonstrate that the results
shown here are independent of this choice of calibrating to a `r 100*(1-reduction)`%
reduction, we conduct the same analysis with a 10% and 30% reduction rate as well.
The corresponding figures can be found in Appendix A.

@@

Figure 3 illustrates how the different forms of adjustment cost can
impact the dynamics of the optimal harvest policy. Corresponding stock
sizes can be seen in Appendix A.  Each panel is generated against the
same sequence of environmental variability so that they can be compared
directly.  The harvest policy chosen shows a systematic deviations from
the adjustment cost free optimum, depending on the structure of the cost function.  
In each case, the optimal solution without any adjustment cost is 
shown by the dashed grey line, with
the policy induced by optimization under the given cost structure 
(equivalent to a `r 100*reduction`% reduction in maximum expected NPV)
overlaid in solid blue.

@@

Figure 4 shows these qualitative patterns still hold at when the models are calibrated 
to a range of different reductions in NPV, from 0% to 30%, though naturally the magnitude
of the differences is greatest when the penalty is larger. Note that results shown in Appendix A provide
further evidence that these patterns are independent of the exact reduction shown by providing
results analogous to Figure 3 (A8, A9) and Figure 5 (A10, A11), and Figure 6 (A6).

@@

```{r}
fig5_caption <- paste("Distribution of revenues from fishing and costs paid to adjustment. Plot shows distributions for the linear costs structure ($\\Pi_1$) only, alternate penalties can be found in the Appendix A. Dockside revenues are always higher than adjustment costs.  When those costs are not accounted for in the policy ('no adjustment costs'), it is possible to obtain only marginally higher revenues, but pay higher adjustment costs.")
```

@@

```{r Figure_6, fig.cap="Relative cost of ignoring adjustment costs when they are present ('ignoring', red) vs assuming the adjustment costs when they are absent ('assuming', blue).  All values are relative to the cost free adjustment optimum, $NPV_0$.  In each case, it is less costly to assume adjustment costs are present when they are not than it is to ignore them when they are present. Appendix A, Figure A6 shows how the magnitude of this effect changes with the size of the reduction to $NPV$.", dev=c("pdf", "svg", "png"), fig.width=7, cache=FALSE}

@@

Figure 6 shows that across each penalty form $\Pi_1$, $\Pi_2$ & $\Pi_3$,
the economic impact of ignoring adjustment costs when they are present
is higher than accounting for them when they are absent. This holds even
after accounting for effect described above: the red bars are consistently
well below the `r (1-reduction)*100`% of $NPV_0$, while the blue bars
are barely below $NPV_0$ (a value of 1 on the vertical axis).  We also
see that the impacts on net present value of the fishery of incorrectly
assuming or ignoring policy adjustment costs are most severe when these
costs are assumed to scale quadratically with the size of the policy
change being implemented ($\Pi_2$).  In Appendix A we
show that pattern holds across different severity of adjustment costs
(robust to choice of $c_i$), though the larger the adjustment cost,
the greater the difference between these two errors.

Note that this analysis does not consider the possibility that an
adjustment cost is present, but of a different functional form than the
policy assumes.  The consequences of managing under the wrong functional
form can be as or more severe than simply ignoring adjustment costs in
the optimization altogether.  We illustrate several of these mismatch
errors in Appendix A.

@@

The biggest differences between the representations of policy adjustment
cost that we consider are between those that have a smoothing effect on
annual quotas and those that do not. The few past studies that incorporate
policy adjustment costs in models of fisheries or other environmental
management contexts [@Feichtinger1994; @Wirl1999] have assumed costs of
policy adjustment increased as a quadratic function of the magnitude of
the policy change being proposed (analogous to $\Pi_2$ formulation). In
effect, this means that small changes to annual quotas incur little
extra cost, but large changes to annual quotas become disproportionately
expensive to make. Including costs of this form smooths inter-annual
variation in the recommended catch quotas. Also, the catches
and remaining stock sizes corresponding to optimal management become more
autocorrelated in time, because it takes the fishery multiple timesteps to harvest
down peaks in abundance that follow large recruitment pulses. The effects
of smoothing here are similar to those predicted when assuming the cost
per unit effort is increasing in the amount of effort expended [@Brown1974;
@Lewis1981; @McGough2009]; (see Appendix A for a summary of the
relevant results) as opposed to associating extra costs with changes to
policy *per se*. Smoothing effects of this type are what one might have
expected when including policy adjustment costs.

@@

Assuming an adjustment cost exists when in fact it is absent is not the
same as using the wrong functional form when it is present.  In Appendix A
we show that choice of the functional form makes a significant
difference.  We show it can be better to ignore policy costs than to derive a 
policy based on assuming the wrong functional form.  This result further
underscores the importance of modelers using caution when seeking to account
for these costs. Assuming an arbitrary form in order to capture the influence
of adjustment costs is thus unlikely to be instructive.

### 2015-08-05 manuscripts/supplement.Rmd
number supplement with A\d not S\d

output:
  pdf_document:
    template: components/elsarticle.latex
    fig_caption: true
    keep_tex: true
---

@@

# Supplementary material #

@@

Figure S1 shows the corresponding stock dynamics for the replicates illustrated in Figure 3 of the main text.  The impact of the adjustment cost on the stock dynamics is less dramatic than it is on the harvest dynamics, due to the stochastic growth in stock sizes and the sequence in which the steps are applied: stock values are assessed without error, then an optimal harvest is determined (using the pre-computed optimal policy based on both the stock assessment and the previous harvest, to account for possible adjustment costs), and then the stock that escapes harvesting recruits stochastically to set the population size for the next year.

Though this results in an effect that is less pronounced, the stock dynamics of Figure S1 reflect the same general trends observed in Main text Figure 3; e.g. the fixed fee $\Pi_3$ results in the most exaggerated fluctuations in the stock sizes.  It is worth noting how changes in policy costs alone, such as introducing or rising fixed costs, can have important knock-on effects to population dynamics such as increased variance, a recent phenomenon observed in important fisheries that has received much attention elsewhere [@Anderson2008].

@@

Figure S2 shows the same summary statistics illustrating the trends
of increasing adjustment costs on the _stock_ dynamics, instead of the
_harvest_ dynamics shown in Figure 4 of the main text.  Note that 
the variance in stock sizes under the quadratic penalty ($\Pi_2$) increases,
while in Figure 4 of the main text the corresponding variance in the
harvest decreases. Under costly adjustment, the undershooting behavior
of the $\Pi_2$ harvest policy cannot absorb as much of the natural 
fluctuation, resulting in increased variation in the stock sizes.

@@

Figure 3 of the main text shows only a single stochastic realization.  Here in
Figure S3 we provide a sample of further realizations (including both
stock and harvest dynamics).

@@

Figure S4 shows that adding quadratic terms to the cost of fishing effort directly, as in @Lewis1981 or @McGough2009, results in a similar smoothing effect as we observe in quadratic adjustment costs ($\Pi_2$), which contrasts sharply with the patterns we observe in other functional forms.  This case corresponds to a profit function defined as:

@@

Figure S5 extends Figure 5 of the main text across all three functional forms of adjustment costs, $\Pi_{1-3}$. To reduce over-plotting of Figure 5 we have also separated the distributions of dockside revenue and adjustment fees by scenario, such that the top row shows the distributions when the policy accounts for adjustment costs, and the bottom shows the policy when it does not (that is, the original @Reed1979 model applied to the scenario of costly adjustment).

@@

Figure S6 extends the results shown in Figure 6 across different severities of adjustment cost, totaling from 10% up to 25% of the $NPV$ of the stock. While the same qualitative pattern is observed in all cases, it is strongest in the higher penalties.

@@

When policy costs are present, assuming the wrong functional form may or
may not be worse than ignoring the presence of policy costs all together.
Figure S7 considers several such cases.  The X axis labels indicate
first which policy is being used to drive the harvest decisions, and
then which penalty is actually being applied.  Thus $\Pi_1\_\Pi_2$ indicates
a policy calculated under the linear ($\Pi_1$) costs, when the reality
involves quadratic ($\Pi_2$) costs.  Though the coefficients of the ($\Pi_2$)
costs have been scaled such that the net effect is the same, the inferred
policies differ.  As a result, the value is slightly more than half of
the optimal free solution when the expected reduction should only be 10% (panel A)
less than optimal free based on the $c_i$ calibration. In this case, ignoring the costs all together
is still worse than using $\Pi_1$ costs -- the red bar is lower than the blue.

@@

These patterns hold across the different size reductions shown (from 10%
to 25%, Figure S8 panels a-d respectively), though net present value becomes negative under the quadratic
form for the larger penalties as costs paid for adjustments exceed
profits derived from harvest.  (Though optimal solutions will always
avoid negative net present values, these scenarios in which policies are
applied under conditions other than assumed in the optimality calculation
have no such guarantee.) For each scenario, results are averaged over
100 replicates.

@@

Figures S8-S11 illustrate that the 
these qualitative patterns still hold at when the models are calibrated 
to a reduction in NPV of 10% or 30%, though naturally the magnitude
of the pattern is greater when the penalty is larger (e.g. 30% vs 10%).

>>>>

output:
  pdf_document:
    template: components/elsarticle.latex
    fig_caption: true
    keep_tex: true
  html_document: 
    toc: true
    toc_depth: 2
    number_sections: true
---

@@

# Appendix A #

@@

Figure A1 shows the corresponding stock dynamics for the replicates illustrated in Figure 3 of the main text.  The impact of the adjustment cost on the stock dynamics is less dramatic than it is on the harvest dynamics, due to the stochastic growth in stock sizes and the sequence in which the steps are applied: stock values are assessed without error, then an optimal harvest is determined (using the pre-computed optimal policy based on both the stock assessment and the previous harvest, to account for possible adjustment costs), and then the stock that escapes harvesting recruits stochastically to set the population size for the next year.

Though this results in an effect that is less pronounced, the stock dynamics of Figure A1 reflect the same general trends observed in Main text Figure 3; e.g. the fixed fee $\Pi_3$ results in the most exaggerated fluctuations in the stock sizes.  It is worth noting how changes in policy costs alone, such as introducing or rising fixed costs, can have important knock-on effects to population dynamics such as increased variance, a recent phenomenon observed in important fisheries that has received much attention elsewhere [@Anderson2008].

@@

Figure A2 shows the same summary statistics illustrating the trends
of increasing adjustment costs on the _stock_ dynamics, instead of the
_harvest_ dynamics shown in Figure 4 of the main text.  Note that 
the variance in stock sizes under the quadratic penalty ($\Pi_2$) increases,
while in Figure 4 of the main text the corresponding variance in the
harvest decreases. Under costly adjustment, the undershooting behavior
of the $\Pi_2$ harvest policy cannot absorb as much of the natural 
fluctuation, resulting in increased variation in the stock sizes.

@@

Figure 3 of the main text shows only a single stochastic realization.  Here in
Figure A3 we provide a sample of further realizations (including both
stock and harvest dynamics).

@@

Figure A4 shows that adding quadratic terms to the cost of fishing effort directly, as in @Lewis1981 or @McGough2009, results in a similar smoothing effect as we observe in quadratic adjustment costs ($\Pi_2$), which contrasts sharply with the patterns we observe in other functional forms.  This case corresponds to a profit function defined as:

@@

Figure A5 extends Figure 5 of the main text across all three functional forms of adjustment costs, $\Pi_{1-3}$. To reduce over-plotting of Figure 5 we have also separated the distributions of dockside revenue and adjustment fees by scenario, such that the top row shows the distributions when the policy accounts for adjustment costs, and the bottom shows the policy when it does not (that is, the original @Reed1979 model applied to the scenario of costly adjustment).

@@

Figure A6 extends the results shown in Figure 6 across different severities of adjustment cost, totaling from 10% up to 25% of the $NPV$ of the stock. While the same qualitative pattern is observed in all cases, it is strongest in the higher penalties.

@@

When policy costs are present, assuming the wrong functional form may or
may not be worse than ignoring the presence of policy costs all together.
Figure A7 considers several such cases.  The X axis labels indicate
first which policy is being used to drive the harvest decisions, and
then which penalty is actually being applied.  Thus $\Pi_1\_\Pi_2$ indicates
a policy calculated under the linear ($\Pi_1$) costs, when the reality
involves quadratic ($\Pi_2$) costs.  Though the coefficients of the ($\Pi_2$)
costs have been scaled such that the net effect is the same, the inferred
policies differ.  As a result, the value is slightly more than half of
the optimal free solution when the expected reduction should only be 10% (panel A)
less than optimal free based on the $c_i$ calibration. In this case, ignoring the costs all together
is still worse than using $\Pi_1$ costs -- the red bar is lower than the blue.

@@

These patterns hold across the different size reductions shown (from 10%
to 25%, Figure A8 panels a-d respectively), though net present value becomes negative under the quadratic
form for the larger penalties as costs paid for adjustments exceed
profits derived from harvest.  (Though optimal solutions will always
avoid negative net present values, these scenarios in which policies are
applied under conditions other than assumed in the optimality calculation
have no such guarantee.) For each scenario, results are averaged over
100 replicates.

@@

Figures A8-A11 illustrate that the 
these qualitative patterns still hold at when the models are calibrated 
to a reduction in NPV of 10% or 30%, though naturally the magnitude
of the pattern is greater when the penalty is larger (e.g. 30% vs 10%).
